---
title: "Questionnaire Analysis"
author: "Or Duek & Tobias R. Spiller"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes

---


```{r}
## Use fear/ emotional pain as a scale to network analysis with 8-factors (also with 7 factors)
## remove item 11 from factors (when adding q42-q44)
## subgroup of high fear (more affected by fear) --> network analysis of PCL-5 without item 11 and with 42-44.
## subgroup of emotional pain --> network analysis of PCL-5 without item 11 and with 42-44.
# load libraries
require(tidyverse)
require(dplyr)
require(ggplot2)
require(bootnet)
require(qgraph)
require(NetworkComparisonTest)
# map
require(sf)
require(mapview)

# tables
require(gtsummary)

#cluster
require(psych)
library(cluster)
library(factoextra)

```

```{r import data, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
# read data
df <- read.csv("Prolific_trauma-exp_only_PCL-V_Randomization.csv",
               na.strings=c("","NA"))

# loading non exposed data
df_open <- read.csv("Prolific _open_to_all_ PCL-V_Randomization.csv",
               na.strings=c("","NA")) 
# keeping just exosed 
df_open <- df_open %>% filter(is.na(trauma_exposure))
```

```{r rename, echo=FALSE}
# Rename variables and convert variable types
df <- df %>% 
  rename(
    sex = Q34,
    education = Q32,
    age = Q33,
    dich = Q47,
    fear = Q39_1,
    pain = Q40_1) %>%
  mutate(
    sex = factor(sex, levels = 1:4, labels = c("Male", "Female", "Non-Binary/Other", "Prefer Not to say")),
    education = factor(education, levels = 1:7, labels = c(
      "lessHighSchool", "HighSchool_GED", "College_nodegree",
      "technical_degree", "BA", "MA/PhD", "Prefer_not")),
    fear = as.numeric(fear),
    pain = as.numeric(pain),
    dich = factor(dich, labels= c("Emotional Pain", "Fearful Reaction")))

# do the same to df_open (a different variable numbers)
df_open <- df_open %>% 
  rename(
    sex = Q34,
    education = Q32,
    age = Q33,
    dich = Q47,
    fear = Q39_1,
    pain = Q40_1) %>%
  mutate(
    sex = factor(sex, levels = 1:4, labels = c("Male", "Female", "Non-Binary/Other", "Prefer Not to say")),
    education = factor(education, levels = 1:7, labels = c(
      "lessHighSchool", "HighSchool_GED", "College_nodegree",
      "technical_degree", "BA", "MA/PhD", "Prefer_not")),
    fear = as.numeric(fear),
    pain = as.numeric(pain),
    dich = factor(dich, labels= c("Emotional Pain", "Fearful Reaction")))

df_open[1,20] <- 32
df_open$age <- as.numeric(df_open$age)
```


```{r summary, echo=FALSE}
# Summarize sex and education across the two datasets
df %>% select(sex, education, dich) %>% rbind(df_open %>% select(sex, education, dich)) %>% tbl_summary(label= list(sex="Sex", education = "Education",
                                                                                                                    dich = "Central Element"))
# Plot age
df_combined <- df %>% select(ResponseId, age) %>%
  rbind(df_open %>% select(ResponseId, age))

df_combined %>%
  ggplot(aes(x = age)) +
  geom_histogram(bins = 10, fill = 'lightblue') +
  theme_minimal() +
  ggtitle("Age Histogram")

```

27% are more centred in fear and the others are more emotional pain

# Network Analysis
```{r include=FALSE}
df3 <- df %>% 
  slice(-1) %>%
  mutate(across(Q2:Q21, as.numeric),
         across(Q50_1:Q50_10, as.numeric),
         across(Q37_1:Q37_9, as.numeric),
         across(Q38_1:Q38_7, as.numeric),
         across(Q51_1:Q51_16, as.numeric)) %>% 
  rowwise(ResponseId) %>%
  mutate(totalPCL = sum(c(Q2, Q3, Q4, Q5, Q6, Q7, Q8, Q9, Q10, Q11, Q12, Q13, Q14, Q15, Q16, Q17, Q18, Q19, Q20, Q21)),
         totalRTQ = sum(c(Q50_1,Q50_2, Q50_3, Q50_3, Q50_5, Q50_6, Q50_7, Q50_8, Q50_9, Q50_10)),
         totalPHQ = sum(Q37_1, Q37_2, Q37_3, Q37_4, Q37_5, Q37_6, Q37_7, Q37_8, Q37_9),
         totalGAD = sum(Q38_1, Q38_2, Q38_3, Q38_4, Q38_5, Q38_6, Q38_7),
         totalASI = sum(Q51_1, Q51_2, Q51_3, Q51_4, Q51_5, Q51_6, Q51_7, Q51_8, Q51_9, Q51_10, Q51_11, Q51_12, Q51_13, Q51_14, Q51_15, Q51_16))

df3 <- df3 %>% 
  ungroup



# the same for df_open
df3_open <- df_open %>% 
  slice(-1) %>%
  mutate(across(c(Q2.1, Q3.1, Q4.1, Q5, Q6.1, Q7.1:Q21), as.numeric),
         across(Q50_1:Q50_10, as.numeric),
         across(Q37_1:Q37_9, as.numeric),
         across(Q38_1:Q38_7, as.numeric),
         across(Q51_1:Q51_16, as.numeric)) %>% 
  rowwise(ResponseId) %>%
  mutate(totalPCL = sum(c(Q2.1, Q3.1, Q4.1, Q5, Q6.1, Q7.1, Q8, Q9, Q10, Q11, Q12, Q13, Q14, Q15, Q16, Q17, Q18, Q19, Q20, Q21)),
         totalRTQ = sum(c(Q50_1,Q50_2, Q50_3, Q50_3, Q50_5, Q50_6, Q50_7, Q50_8, Q50_9, Q50_10)),
         totalPHQ = sum(Q37_1, Q37_2, Q37_3, Q37_4, Q37_5, Q37_6, Q37_7, Q37_8, Q37_9),
         totalGAD = sum(Q38_1, Q38_2, Q38_3, Q38_4, Q38_5, Q38_6, Q38_7),
         totalASI = sum(Q51_1, Q51_2, Q51_3, Q51_4, Q51_5, Q51_6, Q51_7, Q51_8, Q51_9, Q51_10, Q51_11, Q51_12, Q51_13, Q51_14, Q51_15, Q51_16))

df3_open <- df3_open %>%  ungroup

# adding 28 cutoff
df_factor <- df3 %>% #filter(totalPCL>28) %>%
    rename(PCL1 = Q2,
         PCL2 = Q3,
         PCL3 = Q4,
         PCL4 = Q5,
         PCL5 = Q6,
         PCL6 = Q7,
         PCL7 = Q8,
         PCL8 = Q9,
         PCL9 = Q10,
         PCL10 = Q11,
         PCL11 = Q12,
         PCL12 = Q13,
         PCL13 = Q14,
         PCL14 = Q15,
         PCL15 = Q16,
         PCL16 = Q17,
         PCL17 = Q18,
         PCL18 = Q19,
         PCL19 = Q20,
         PCL20 = Q21)

# the same for df_open
df_factor_open <- df3_open %>% #filter(totalPCL>28) %>%
    rename(PCL1 = Q2.1,
         PCL2 = Q3.1,
         PCL3 = Q4.1,
         PCL4 = Q5,
         PCL5 = Q6.1,
         PCL6 = Q7.1,
         PCL7 = Q8,
         PCL8 = Q9,
         PCL9 = Q10,
         PCL10 = Q11,
         PCL11 = Q12,
         PCL12 = Q13,
         PCL13 = Q14,
         PCL14 = Q15,
         PCL15 = Q16,
         PCL16 = Q17,
         PCL17 = Q18,
         PCL18 = Q19,
         PCL19 = Q20,
         PCL20 = Q21)


```

```{r eval=TRUE, message=TRUE, warning=FALSE, echo=FALSE}
# build cluster scores
df3 <- df_factor %>% mutate(
  RE1 = PCL1 + PCL2 + PCL3,
  RE2 = PCL4 + PCL5,
  Av = PCL6 + PCL7,
  Na = PCL8 + PCL9 + PCL10 + PCL11,
  An = PCL12 + PCL13 + PCL14,
  Eb = PCL15 + PCL16,
  Aa = PCL17 + PCL18,
  Da = PCL19 + PCL20,
  Reexp = RE1 + RE2 # adding for 7-factor model
)


# same for df_open
df3_open <- df_factor_open %>% mutate(
  RE1 = PCL1 + PCL2 + PCL3,
  RE2 = PCL4 + PCL5,
  Av = PCL6 + PCL7,
  Na = PCL8 + PCL9 + PCL10 + PCL11,
  An = PCL12 + PCL13 + PCL14,
  Eb = PCL15 + PCL16,
  Aa = PCL17 + PCL18,
  Da = PCL19 + PCL20,
  Reexp = RE1 + RE2 # adding for 7-factor model
)

# combine the two datasets
df_combined <- df3 %>% select(ResponseId, Reexp, RE1, RE2, Av, Na, An, Eb, Aa, Da, fear, pain, totalRTQ, totalASI, totalGAD, Q42, Q43, Q44, Q45, Q46, dich, starts_with("PCL")) %>%
  rbind(df3_open %>% select(ResponseId, Reexp, RE1, RE2, Av, Na, An, Eb, Aa, Da, fear, pain, totalRTQ, totalASI, totalGAD, Q42, Q43, Q44, Q45, Q46, dich, starts_with("PCL")))
```

//### We have a total of 664 participants beyond the PCL 28 cutoff

## 8-factors with fear and pain
```{r echo=FALSE}
# bootnet
df4 <- df_combined %>% select(RE1, RE2, Av, Na, An, Eb, Aa, Da, fear, pain) 

net <- estimateNetwork(df4, "ggmModSelect")
plot(net)

```



## 7-factors with fear and pain

```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df4.1 <- df_combined %>% select(Reexp, Av, Na, An, Eb, Aa, Da, fear, pain) 

net1 <- estimateNetwork(df4.1, "ggmModSelect")
plot(net1)
```

## 8-factors including Rumination (RTQ) and Anxiety (ASI)

```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df5 <- df_combined %>% select(RE1, RE2, Av, Na, An, Eb, Aa, Da, fear, pain, totalRTQ, totalASI) 

net5 <- estimateNetwork(df5, "ggmModSelect")
plot(net5)
```

## 7-factor model including rumination and anxiety
```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df5.1 <- df_combined %>% select(Reexp, Av, Na, An, Eb, Aa, Da, fear, pain, totalRTQ, totalASI) 

net5.1 <- estimateNetwork(df5.1, "ggmModSelect")
plot(net5.1)
```

## 8-factor model with Rumination and GAD-7
```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df6 <- df_combined %>% select(RE1, RE2, Av, Na, An, Eb, Aa, Da, fear, pain, totalRTQ, totalGAD) 

net6 <- estimateNetwork(df6, "ggmModSelect")
plot(net6)
```

## 7-factor model with Rumination and GAD-7
```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df6.1 <- df_combined %>% select(Reexp, Av, Na, An, Eb, Aa, Da, fear, pain, totalRTQ, totalGAD) 

net6.1 <- estimateNetwork(df6.1, "ggmModSelect")
plot(net6.1)
```

## Network Analysis of 8-Factor just for Fear
```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df7 <- df_combined %>% filter(dich=="Fearful Reaction") %>% select(RE1, RE2, Av, Na, An, Eb, Aa, Da) 

net7 <- estimateNetwork(df7, "ggmModSelect")
plot(net7)
```

## Now just pain
```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df7.1 <- df_combined %>% filter(dich=="Emotional Pain") %>% select(RE1, RE2, Av, Na, An, Eb, Aa, Da) 

net7.1 <- estimateNetwork(df7.1, "ggmModSelect")
plot(net7.1)
```

## Compare networks
```{r}

comp1 <- NCT(net7, net7.1, test.edges = TRUE, verbose = FALSE)

plot(comp1, what = "network")
```

Some differences between edges of the network (mostly re1, re2 with NA, Av and Da), but the network itself is not significantly different.

## Compare networks with the 7-factor model

### Fear
```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df8 <- df_combined %>% filter(dich=="Fearful Reaction") %>% select(Reexp, Av, Na, An, Eb, Aa, Da) 

net8 <- estimateNetwork(df8, "ggmModSelect")
plot(net8)
```

### Pain
```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df8.1 <- df_combined %>% filter(dich=="Emotional Pain") %>% select(Reexp, Av, Na, An, Eb, Aa, Da) 

net8.1 <- estimateNetwork(df8.1, "ggmModSelect")
plot(net8.1)
```

### Compare networks (7-factors)
```{r echo=FALSE, message=FALSE, warning=FALSE}

comp2 <- NCT(net8, net8.1, test.edges = TRUE, verbose = FALSE)

plot(comp2, what = "network")
```

Results are similar


# Using three questions instead of PCL-11 and two instead of PCL-10

Q42 = Fear and Horror
Q43 = Guilt and Shame
Q44 = Anger
Q45 - Blaming yourself
Q46 - Blaming someone else

```{r echo=FALSE, message=FALSE, warning=FALSE}
# generate network of just PCL, but without item -11

df_new <- df_combined %>%   
  select(PCL1:PCL20, Q42, Q43, Q44, -PCL11) 

net_PCL_new_items <- estimateNetwork(df_new, "ggmModSelect")
plot(net_PCL_new_items)#, layout = "circle")
```



## Build 8-factor without PCL11 and with the three items

```{r eval=TRUE, message=TRUE, warning=FALSE, echo=FALSE}
# build cluster scores
df3_new <- df_combined %>% mutate(
  RE1 = PCL1 + PCL2 + PCL3,
  RE2 = PCL4 + PCL5,
  Av = PCL6 + PCL7,
  Na = PCL8 + PCL9 + PCL10,
  An = PCL12 + PCL13 + PCL14,
  Eb = PCL15 + PCL16,
  Aa = PCL17 + PCL18,
  Da = PCL19 + PCL20,
  Reexp = RE1 + RE2 # adding for 7-factor model
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df4_new <- df3_new %>% select(RE1, RE2, Av, Na, An, Eb, Aa, Da, Q42, Q43, Q44) 

net4_new <- estimateNetwork(df4_new, "ggmModSelect")
plot(net4_new)
```



## Build 7-factor without PCL11 and with the three items


```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df4.1_new <- df3_new %>% select(Reexp, Av, Na, An, Eb, Aa, Da, Q42, Q43, Q44) 

net4.1_new <- estimateNetwork(df4.1_new, "ggmModSelect")
plot(net4.1_new)
```

## Run similar analysis but on the different groups

### Emotional pain

```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df10 <- df3_new %>% filter(dich=="Emotional Pain") %>% select(Reexp, Av, Na, An, Eb, Aa, Da, Q42, Q43, Q44, totalRTQ) 

net10 <- estimateNetwork(df10, "ggmModSelect")
plot(net10)
```

### Fearful reaction

```{r echo=FALSE, message=FALSE, warning=FALSE}
# bootnet
df10.1 <- df3_new %>% filter(dich=="Fearful Reaction") %>% select(Reexp, Av, Na, An, Eb, Aa, Da, Q42, Q43, Q44, totalRTQ) 

net10.1 <- estimateNetwork(df10.1, "ggmModSelect")
plot(net10.1)
```

## Bar Plot of each symptom average score for the Fear and Emotional Pain groups

```{r, fig.width=12, fig.height=6}

# generate barplot of average score for each symptom in the fear/emotional groups
df_new1 <- df_combined %>% select(PCL1:PCL20, Q42, Q43, Q44, dich)
df_new1 %>% drop_na() %>% group_by(dich) %>% summarise_all(mean) 

df_new1 %>%
  drop_na() %>%
  group_by(dich) %>%
  summarise_all(mean, na.rm = TRUE) %>%
  pivot_longer(-dich, names_to = "Item", values_to = "Average") %>%
  mutate(Item = factor(Item, levels = paste0('PCL', 1:20))) %>%
  ggplot(aes(x = Item, y = Average, fill = dich)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Category", y = "Average Score", title = "Average Scores for Each Item") +
  theme_minimal()

```



# Factor analysis

Exploratory Factor (EFA)
Assess whether the new emotion items align with the DSM-5 model by conducting EFA assuming a 4 factor structure.
```{r}
# build correlation matrix of the data
cor_matrix <- df_combined %>% select(PCL1:PCL20, Q42:Q46) %>% drop_na() %>% cor()

# Perform EFA using principal axis factoring (PAF) with oblimin rotation
efa_result <- fa(cor_matrix, nfactors = 4, rotate = "oblimin", fm = "pa", 
                 missing = "pairwise", scores = "regression")

# Print the factor loadings
print(efa_result$loadings, digits = 2, cutoff = 0.4)
```
