---
title: "Questionnaire Analysis"
author: "Or Duek & Tobias R. Spiller"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes

---


```{r, results='hide',  message=FALSE, warning=FALSE, include=FALSE}
# load libraries
require(tidyverse)
require(dplyr)
require(ggplot2)
require(bootnet)
require(qgraph)
# map
require(sf)
require(mapview)

# tables
require(gtsummary)
```

```{r, results='hide', message=FALSE, warning=FALSE, include=FALSE}
# load data - removing first line (which is dictionary)
df_all <- read.csv('PCL-V+Randomization_first_batch_8_17_22.csv', header = F)

# grab headers from second line
headers <- df_all[2,]
# grab all data
df <- df_all[3:nrow(df_all),]
# make DF great again
colnames(df) = headers
summary(data.frame(df))

# save to csv
write_csv(df, "pcl.csv")
# now read it again
df <- read.csv('pcl.csv')
# for reference - grab first line (dictionary)
dict <- df_all[1,]
# print if needed
#as.character(dict)
```



# Location of responders

```{r}
df %>% dplyr::filter(Q62==2) %>% mapview(xcol = "LocationLongitude", ycol = "LocationLatitude", crs = 4269,  grid = F)
```


Two people marked that they live in US but they actually filled it from a different location

## Attention Checks

Let's see how many failed the attention check

- Asking about Moon and Earth
- Asking about colors (say all)
- Asking about No. of traumas (say 5)

```{r}

df_attent <- df %>% select(Q66, Q65, Q64)

# moon
df_attent %>% drop_na(Q66) %>% select(Q66) %>% tbl_summary()

# traumas
df_attent %>% drop_na(Q64) %>% select(Q64) %>% tbl_summary()

# colors (I remove based on 64, as 65 is character, so no NAs)
df_attent %>% drop_na(Q64) %>% select(Q65) %>% tbl_summary()

```


Lets just grab the dichotomous question (which is more central emotional pain or fear), and see if it correlates with the manual scale

```{r}
# 40  -dicho, 67 - trauma reminder, 68 - emotional pain

df2 <- df[2:nrow(df),c(9,40, 67, 68)]
# changing column names
colnames(df2)[2] = "dich"
colnames(df2)[3] = "fear"
colnames(df2)[4] = "pain"

#df2[,2] <- as.numeric(df2[,2])
df2[,3] <- as.numeric(df2[,3])
df2[,4] <- as.numeric(df2[,4])
summary(df2)
df2 <- drop_na(df2) # 18 subjects

ggplot(df2, aes(x=dich, y=pain, group = dich)) + geom_boxplot() + theme_minimal() + ggtitle("Emotional pain as factor of the Dichotomous question")

ggplot(df2, aes(x=dich, y=fear, group = dich)) + geom_boxplot() + theme_minimal() + ggtitle("Fear as factor of the Dichotomous question")
```


OK, looks relatively ok - with each group more relating to the relevant scale

## Network Analysis


```{r}
# organize dataframe including PCL and RTQ


df_pcl <- df[,41:60]
df_rtq <- df[,86:95]
df_pcl_rtq <- drop_na(cbind(df_pcl, df_rtq))

```

Some sanity checks

```{r}
# rtq total
hist(rowSums(df_pcl_rtq[,21:30]))

# pcl total
hist(rowSums(df_pcl_rtq[,1:20]))

```


Building network

```{r}
n1 <- estimateNetwork(df_pcl_rtq[,1:20], default="huge", corMethod = "cor", corArgs = list(method="spearman"), threshold=FALSE)
g2 <- plot(n1, legend.cex=.5, vsize=7)
```

