---
title: "Questionnaire Analysis"
author: "Or Duek & Tobias R. Spiller"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes

---


```{r, results='hide',  message=FALSE, warning=FALSE, include=FALSE}
# load libraries
require(tidyverse)
require(dplyr)
require(ggplot2)
require(bootnet)
require(qgraph)
# map
require(sf)
require(mapview)

# tables
require(gtsummary)
```

```{r, results='hide', message=FALSE, warning=FALSE, include=FALSE}
# load data - removing first line (which is dictionary)
#df_all <- read.csv('PCL-V+Randomization_first_batch_8_17_22.csv', header = F)
df_all <- read.csv('data/PCL-V+Randomization+&+NAPS+Image+Rating_November+4,+2022_10.43_clean.csv')
# grab headers from second line
#headers <- df_all[1,]
# grab all data
df <- df_all#[3:nrow(df_all),]
# make DF great again
c#olnames(df) = headers
#summary(data.frame(df))

# save to csv
# write_csv(df, "data/pcl_clean.csv")
# # now read it again
# df <- read.csv('data/pcl_clean.csv')
# # for reference - grab first line (dictionary)
# dict <- df_all[1,]
# print if needed
#as.character(dict)
```

# Basic Demographics

- Age and Sex

```{r}
#
df %>%  select(Q34) %>% mutate(Q34 = factor(Q34, levels=c(1:4), labels = c("Male","Female","Non-Binary/Other","Prefer Not to say"))) %>%
  tbl_summary(label = list(Q34 ~ "Sex")) # need to add labels

ggplot(df, aes(x=Q33)) + geom_histogram(bins=10, fill = 'lightblue') +theme_minimal() + ggtitle("Age Histogram (Before omitting")
```


# Location of responders

```{r}
df %>% dplyr::filter(Q62==2) %>% drop_na(LocationLongitude) %>% mapview(xcol = "LocationLongitude", ycol = "LocationLatitude", crs = 4269,  grid = F)
```


Two people marked that they live in US but they actually filled it from a different location

# Attention Checks

Let's see how many failed the attention check

- Asking about Moon and Earth
- Asking about colors (say all)
- Asking about No. of traumas (say 5)

```{r}

df_attent <- df %>% select(Q66, Q65, Q64)

# moon
df_attent %>% drop_na(Q66) %>% select(Q66) %>% tbl_summary()

# traumas
df_attent %>% drop_na(Q64) %>% select(Q64) %>% tbl_summary()

# colors (I remove based on 64, as 65 is character, so no NAs)
df_attent %>% drop_na(Q64) %>% select(Q65) %>% tbl_summary()

```

# Sanity 

Lets just grab the dichotomous question (which is more central emotional pain or fear), and see if it correlates with the manual scale

```{r}
# 40  -dicho (1=pain, 2=fear), 67 - trauma reminder, 68 - emotional pain

df2 <- df %>% select(Q47, Q39_1, Q40_1)#$[2:nrow(df),c(9,40, 67, 68)]
# changing column names
colnames(df2)[1] = "dich"
colnames(df2)[2] = "fear"
colnames(df2)[3] = "pain"
df2$dich <- factor(df2$dich)
summary(df2)
df2 <- drop_na(df2) # 44 subjects

ggplot(df2, aes(x=dich, y=pain, group = dich)) + geom_boxplot() + theme_minimal() + ggtitle("Emotional pain as factor of the Dichotomous question")

ggplot(df2, aes(x=dich, y=fear, group = dich)) + geom_boxplot() + theme_minimal() + ggtitle("Fear as factor of the Dichotomous question")
```


OK, looks relatively ok - with each group more relating to the relevant scale

# Histograms

## Hisotgams of RTQ and PCL

```{r}
# organize dataframe including PCL and RTQ
df_pcl <- df[,41:60]
df_rtq <- df %>% select(starts_with('Q50'))#df[,86:95]
df_pcl_rtq <- drop_na(cbind(df_pcl, df_rtq))
```

Some sanity checks

```{r}
# rtq total
hist(rowSums(df_pcl_rtq[,21:30]))

# pcl total
hist(rowSums(df_pcl_rtq[,1:20]))

summary(rowSums(df_pcl_rtq[,1:20]))
```


<!-- ## Building network -->

<!-- ```{r} -->
<!-- n1 <- estimateNetwork(df_pcl_rtq[,1:20], default="huge", corMethod = "cor", corArgs = list(method="spearman"), threshold=FALSE) -->
<!-- g2 <- plot(n1, legend.cex=.5, vsize=7) -->

### Histogram of each PCL item

```{r}
for (i in 1:20) {
  p <- ggplot(data=df_pcl_rtq, aes(x=df_pcl_rtq[,i])) + geom_histogram(binwidth = 1) + theme_bw() + ggtitle('PCL', i)
  #hist(df_pcl_rtq[,i])
  print(p)
}



```

