---
title: "Questionnaire Analysis"
author: "Or Duek & Tobias R. Spiller"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes

---


```{r, results='hide',  message=FALSE, warning=FALSE, include=FALSE}
# load libraries
require(tidyverse)
require(dplyr)
require(ggplot2)
require(bootnet)
require(qgraph)
# map
require(sf)
require(mapview)

# tables
require(gtsummary)
# xls
library(readxl)
```

# Analysing Prolific Data

```{r}
# read data
df <- read.csv('Prolific_trauma-exp_only_PCL-V_Randomization.csv', na.strings=c("","NA"))
```

# Basic Demographics

- Age and Sex

```{r}
#
df %>%  select(Q34) %>% mutate(Q34 = factor(Q34, levels=c(1:4), labels = c("Male","Female","Non-Binary/Other","Prefer Not to say"))) %>%
  tbl_summary(label = list(Q34 ~ "Sex")) # need to add labels

# age is not clean, need to go over, if not numeric then set as NA
df <- df %>% mutate (
  age = as.numeric(df$Q33)
)

ggplot(df, aes(x=age)) + geom_histogram(bins=10, fill = 'lightblue') +theme_minimal() + ggtitle("Age Histogram (Before omitting")

# Education
df %>%  select(Q32) %>% mutate(Q32 = factor(Q32, levels=c(1:6), labels = c("1","2","3","4", "5","6"))) %>%
  tbl_summary(label = list(Q32 ~ "Education")) # need to add labels
```

# Location of responders

```{r}
df %>%  drop_na(LocationLongitude) %>% mapview(xcol = "LocationLongitude", ycol = "LocationLatitude", crs = 4269,  grid = F)
```

# Attention Checks

Let's see how many failed the attention check

- Asking about Moon and Earth
- Asking about colors (say all)
- Asking about No. of traumas (say 5)

```{r}

df %>% select(Q65) %>% drop_na()  %>% tbl_summary()

```

Only one person failed attention check


# Sanity 

Lets just grab the dichotomous question (which is more central emotional pain or fear), and see if it correlates with the manual scale

```{r}
# 40  -dicho (1=pain, 2=fear), 67 - trauma reminder, 68 - emotional pain

df2 <- df %>% select(Q47, Q39_1, Q40_1)#$[2:nrow(df),c(9,40, 67, 68)]
# changing column names
colnames(df2)[1] = "dich"
colnames(df2)[2] = "fear"
colnames(df2)[3] = "pain"
df2$dich <- factor(df2$dich)
summary(df2)
df2 <- drop_na(df2) # 44 subjects

ggplot(df2, aes(x=dich, y=pain, group = dich)) + geom_boxplot() + theme_minimal() + ggtitle("Emotional pain as factor of the Dichotomous question")

ggplot(df2, aes(x=dich, y=fear, group = dich)) + geom_boxplot() + theme_minimal() + ggtitle("Fear as factor of the Dichotomous question")
summary(lm(fear ~ dich, data=df2))
```
## Calculate Total Scores

```{r}
# PCL items: Q2 - Q21
# PCL break: Q42 (fear/horror), Q43 (guilt/shame), Q44 (anger), Q45(blame yourself), Q46(Blame others)
# These will replace Q11 and Q12
# RTQ items: Q50_1 - Q50_10
# ASI items: Q51_1 - Q51_16
# GAD items: Q38_1 - Q38_7
# PHQ-9: Q37_1 - Q37_9
```

### PCL-5 Total Score
```{r}
df <- df %>% rowwise(ResponseId) %>% mutate(totalPCL = sum(c(Q2, Q3, Q4, Q5, Q6, Q7, Q8, Q9, Q10, Q11, Q12, Q13, Q14, Q15, Q16, Q17, Q18, Q19, Q20, Q21)),
                                            totalRTQ = sum(c(Q50_1,Q50_2, Q50_3, Q50_3, Q50_5, Q50_6, Q50_7, Q50_8, Q50_9, Q50_10))
                                                                                        )
hist(df$totalPCL)
hist(df$totalRTQ)
```

### Test histogram of RTQ within different cutoffs of PCL
```{r}

df_highP <- df %>% filter(totalPCL>=28)
df_lowP <- df %>% filter(totalPCL<28)

par(mfrow=c(2,1))
 
#Make the plot
par(mar=c(2,5,1,2))
hist(df_highP$totalRTQ , main="" , ylab="Frequency for high PCL", xlab="", xaxt="n", ylim=c(0,100), las=1 , col="slateblue1", breaks=10)
hist(df_lowP$totalRTQ, main="", ylab="Frequency for low PCL", xlab="RTQ-10 Score", ylim=c(100,0),las=1 , col="tomato3"  , breaks=10)

```

### Correlation between RTQ and PCL
```{r}
df <- df %>% mutate(
  pcl_H_L = case_when(totalPCL>=28 ~ "HighPCL",
                      totalPCL<28 ~ "LowPCL")
)
df %>% drop_na(totalPCL) %>% ggplot(aes(x=totalRTQ, y=totalPCL)) + geom_point(aes(color=pcl_H_L)) + geom_smooth(method='lm') + theme_minimal()

cor.test(df$totalPCL, df$totalRTQ)
```

Correlation of 0.66.

