---
title: "Analysis of Prolific Data for B & I"
author: "Or Duek & Tobias R. Spiller"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes

---
# Introduction & Basic Analysis

The following code performs basic data analysis of a questionnaire dataset collected through the Prolific platform. The dataset includes demographic information as well as responses to various psychological measures, including the Posttraumatic Stress Disorder Checklist for DSM-5 (PCL-5), the Revised Trauma Questionnaire (RTQ), the Patient Health Questionnaire (PHQ), the Generalized Anxiety Disorder scale (GAD), and the Anxiety Sensitivity Index (ASI). The code explores the demographic characteristics of the sample, the quality of the data (attention checks), and the relationship between different measures using network and exploratory factor analysis.

## Libraries and Data Reading

The first part of the code loads necessary libraries and reads the CSV file with the dataset.
```{r Load libraries, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
# data handling
library(readxl)
library(tidyverse)
library(ggplot2)

# Basic analysis
library(psych)
library(qgraph)
library(cluster)
library(factoextra)

#Networks
library(bootnet)
library(qgraph)
library(psychonetrics)

# map
library(sf)
#require(mapview)

# tables
library(gtsummary)
```

```{r import data, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
# read data
df <- read.csv("Prolific_trauma-exp_only_PCL-V_Randomization.csv",
               na.strings=c("","NA"))
```

## Basic Demographics

The next part of the code explores the basic demographics of the sample, including age, sex, and education level. Age is plotted as a histogram, while sex and education level are summarized using the tbl_summary function from the gtsummary package.
```{r rename, echo=FALSE}
# Rename variables and convert variable types
df <- df %>% 
  rename(
    sex = Q34,
    education = Q32,
    age = Q33,
    dich = Q47,
    fear = Q39_1,
    pain = Q40_1) %>%
  mutate(
    sex = factor(sex, levels = 1:4, labels = c("Male", "Female", "Non-Binary/Other", "Prefer Not to say")),
    education = factor(education, levels = 1:7, labels = c(
      "lessHighSchool", "HighSchool_GED", "College_nodegree",
      "technical_degree", "BA", "MA/PhD", "Prefer_not")),
    fear = as.numeric(fear),
    pain = as.numeric(pain),
    dich = factor(dich))
```

```{r summary, echo=FALSE}
# Summarize sex and education
tbl_summary(df %>% select(sex), label = list(sex = "Sex"))
tbl_summary(df %>% select(education), label = list(education = "Education"))

# Plot age
ggplot(df, aes(x = age)) + 
  geom_histogram(bins = 10, fill = 'lightblue') +
  theme_minimal() +
  ggtitle("Age Histogram")
  ggtitle("Age Histogram")
```

## Location of Responders

The code uses the mapview package to visualize the location of responders on a map. [disabled as I (Tobias) cannot install the package]
```{r eval=FALSE, include=FALSE}
df %>%
  drop_na(LocationLongitude) %>% 
  mapview(xcol = "LocationLongitude", ycol = "LocationLatitude", crs = 4269,  grid = F)
```

## Attention Checks

The code checks the quality of the data by examining the responses to attention checks. The results are summarized using the tbl_summary function from the gtsummary package. Only *one* person failed attention check

Let's see how many failed the attention check
- Asking about Moon and Earth
- Asking about colors (say all)
- Asking about No. of traumas (say 5)
```{r}
df %>% 
  select(Q65) %>% 
  drop_na()  %>% 
  tbl_summary()
```

## Sanity Check

The code performs a sanity check by examining the relationship between a dichotomous question about central emotional pain or fear and a manual scale that measures emotional pain and fear. The results are visualized using boxplots and a linear regression model.
```{r}
# Rename columns and convert variable types
# Plot emotional pain and fear as a factor of the dichotomous question
ggplot(df, aes(x=dich, y=pain, group = dich)) + 
  geom_boxplot() + 
  theme_minimal() + 
  ggtitle("Emotional pain as factor of the Dichotomous question")

ggplot(df, aes(x=dich, y=fear, group = dich)) + 
  geom_boxplot() + 
  theme_minimal() + 
  ggtitle("Fear as factor of the Dichotomous question")

# Summary of linear regression model
summary(lm(fear ~ dich, data=df))
```

## Calculate Total Scores

The code calculates total scores for each psychological measure by summing the item scores for each measure. 
```{r include=FALSE}
df3 <- df %>% 
  slice(-1) %>%
  mutate(across(Q2:Q21, as.numeric),
         across(Q50_1:Q50_10, as.numeric),
         across(Q37_1:Q37_9, as.numeric),
         across(Q38_1:Q38_7, as.numeric),
         across(Q51_1:Q51_16, as.numeric)) %>% 
  rowwise(ResponseId) %>%
  mutate(totalPCL = sum(c(Q2, Q3, Q4, Q5, Q6, Q7, Q8, Q9, Q10, Q11, Q12, Q13, Q14, Q15, Q16, Q17, Q18, Q19, Q20, Q21)),
         totalRTQ = sum(c(Q50_1,Q50_2, Q50_3, Q50_3, Q50_5, Q50_6, Q50_7, Q50_8, Q50_9, Q50_10)),
         totalPHQ = sum(Q37_1, Q37_2, Q37_3, Q37_4, Q37_5, Q37_6, Q37_7, Q37_8, Q37_9),
         totalGAD = sum(Q38_1, Q38_2, Q38_3, Q38_4, Q38_5, Q38_6, Q38_7),
         totalASI = sum(Q51_1, Q51_2, Q51_3, Q51_4, Q51_5, Q51_6, Q51_7, Q51_8, Q51_9, Q51_10, Q51_11, Q51_12, Q51_13, Q51_14, Q51_15, Q51_16))

df3 <- df3 %>% 
  ungroup
```

Histograms are used to visualize the distributions of the total scores for each measure.
```{r echo=FALSE}
# Plot histograms of total scores for each measure
hist(df3$totalPCL, main = "Total PCL Score Distribution")
hist(df3$totalRTQ, main = "Total RTQ Score Distribution")
hist(df3$totalPHQ, main = "Total PHQ Score Distribution")
hist(df3$totalGAD, main = "Total GAD Score Distribution")
hist(df3$totalASI, main = "Total ASI Score Distribution")
```

## Correlation Analysis

The code examines the relationships between different psychological measures using scatterplots and correlation tests. The results are visualized using ggplot2 and tested using the cor.test function.

### Test histogram of RTQ within different cutoffs of PCL
```{r}
df_highP <- df3 %>% filter(totalPCL >= 28)
df_lowP <- df3 %>% filter(totalPCL < 28)

#Make the plot
ggplot() +
  geom_histogram(data = df_highP, aes(x = totalRTQ), binwidth = 2, color = "black", fill = "slateblue1", alpha = 0.7) +
  geom_histogram(data = df_lowP, aes(x = totalRTQ), binwidth = 2, color = "black", fill = "tomato3", alpha = 0.7) +
  labs(x = "", y = "Frequency", title = "Histogram of RTQ within different cutoffs of PCL") +
  scale_y_continuous(limits = c(0, 100)) +
  theme_minimal()
```

### Correlation between RTQ and PCL
```{r warning=FALSE}
df3 %>%
  ggplot(aes(x = totalRTQ, y = totalPCL, color = factor(ifelse(totalPCL >= 28, "HighPCL", "LowPCL")))) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Correlation between RTQ and PCL") +
  theme_minimal()

cor.test(df3$totalPCL, df3$totalRTQ)
```

### Correlation between PCL and GAD
```{r warning=FALSE}
df3 %>%
  ggplot(aes(x = totalGAD, y = totalPCL, color = factor(ifelse(totalPCL >= 28, "HighPCL", "LowPCL")))) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Correlation between PCL and GAD") +
  theme_minimal()

cor.test(df3$totalPCL, df3$totalGAD)
```

### Correlation between PCL and ASI
```{r warning=FALSE}
df3 %>%
  ggplot(aes(x = totalASI, y = totalPCL, color = factor(ifelse(totalPCL >= 28, "HighPCL", "LowPCL")))) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Correlation between PCL and ASI") +
  theme_minimal()

cor.test(df3$totalPCL, df3$totalASI)
```

### Correlation between RTQ and ASI
```{r warning=FALSE}
df3 %>%
  ggplot(aes(x = totalASI, y = totalRTQ, color = factor(ifelse(totalPCL >= 28, "HighPCL", "LowPCL")))) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Correlation between PCL and RTQ") +
  theme_minimal()
```


# OLD LNM ANALYSIS - NOT RUN - COULD BE DELETED 
```{r LNM rename, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
df_factor <- df3 %>% 
    rename(PCL1 = Q2,
         PCL2 = Q3,
         PCL3 = Q4,
         PCL4 = Q5,
         PCL5 = Q6,
         PCL6 = Q7,
         PCL7 = Q8,
         PCL8 = Q9,
         PCL9 = Q10,
         PCL10 = Q11,
         PCL11 = Q12,
         PCL12 = Q13,
         PCL13 = Q14,
         PCL14 = Q15,
         PCL15 = Q16,
         PCL16 = Q17,
         PCL17 = Q18,
         PCL18 = Q19,
         PCL19 = Q20,
         PCL20 = Q21)

```


```{r LNM 3 Factor, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
# Rename
# Factor structure
Latents_3_factor<-c("PCL","RTQ", "ASI")

# Variables
vars_3 <- c("PCL1", "PCL2", "PCL3", "PCL4", "PCL5", "PCL6", "PCL7", "PCL8", "PCL9", "PCL10", 
            "PCL11", "PCL12", "PCL13", "PCL14", "PCL15", "PCL16", "PCL17", "PCL18", "PCL19", "PCL20", 
            "Q50_1", "Q50_2", "Q50_3", "Q50_4", "Q50_5", "Q50_6", "Q50_7", "Q50_8", "Q50_9", "Q50_10", 
            "Q51_1", "Q51_2", "Q51_3", "Q51_4", "Q51_5", "Q51_6", "Q51_7", "Q51_8", "Q51_9", "Q51_10", 
            "Q51_11", "Q51_12", "Q51_13", "Q51_14", "Q51_15", "Q51_16")


# Individual factors
Lambda_3_factor <- matrix(0,46,3)
Lambda_3_factor[1:20,1] <- 1
Lambda_3_factor[21:30,2] <- 1
Lambda_3_factor[31:46,3] <- 1
```

Second, we define the alpha level and set the seed.
```{r LNM set alpha, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
alpha = 0.01
set.seed(NULL)
```

Third, we define the LNMs.
```{r LNM define model, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
# Define models
lnmMod_3_factor <-  lnm(df_factor, vars=vars_3,  lambda = Lambda_3_factor,  latents = Latents_3_factor, identification = "variance")
```

Fourth, we run the LNM.
```{r LNM run model, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
# Run & Remove non-sig latent edges:
lnmMod_3_prune <- lnmMod_3_factor %>% runmodel %>% psychonetrics::prune(alpha = alpha) 
```

Results
Fit measures
```{r LNM fit, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
lnmMod_3_prune@fitmeasures[c("chisq","pvalue","df", "rmsea","cfi", "tli", "aic.ll", "bic")]
```

Plot
```{r LNM Plot, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
qgraph(lnmMod_3_prune@modelmatrices[[1]]$omega_zeta, labels = Latents_3_factor,
       theme = "colorblind", vsize = 10)
```


 8 factor models
```{r LNM 8 Factor, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
# Rename
# Factor structure
Latents_8_factor<-c("Re1","Re2", "Av", "Na", "An","Eb","Aa","Da")

# Variables
vars_8 = c("PCL1", "PCL2", "PCL3","PCL4","PCL5","PCL6","PCL7","PCL8" ,"PCL9",
        "PCL10","PCL11","PCL12","PCL13","PCL14","PCL15","PCL16","PCL17", "PCL18","PCL19","PCL20")

# Individual factors
Lambda_8_factor <- matrix(0,20,8)
Lambda_8_factor[1:3,1] <- 1
Lambda_8_factor[4:5,2] <- 1
Lambda_8_factor[6:7,3] <- 1
Lambda_8_factor[8:11,4] <- 1
Lambda_8_factor[12:14, 5] <- 1
Lambda_8_factor[15:16,6] <- 1
Lambda_8_factor[17:18,7] <- 1
Lambda_8_factor[19:20,8] <- 1
```

7 Factor Model
```{r LNM 7 Factor, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
# Factor structure
Latents_7_factor<-c("Re", "Av", "Na", "An","Eb","Aa","Da")

# Individual factors
Lambda_7_factor <- matrix(0,20,7)
Lambda_7_factor[1:5,1] <- 1
Lambda_7_factor[6:7,2] <- 1
Lambda_7_factor[8:11,3] <- 1
Lambda_7_factor[12:14,4] <- 1
Lambda_7_factor[15:16,5] <- 1
Lambda_7_factor[17:18,6] <- 1
Lambda_7_factor[19:20,7] <- 1
```

4 Factor DSM  Model
```{r LNM 5 Factor, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
# Factor structure
Latents_4_factor<-c("Re", "Av", "Mood", "Arousal")

# Individual factors
Lambda_4_factor <- matrix(0,20,4)
Lambda_4_factor[1:5,1] <- 1
Lambda_4_factor[6:7,2] <- 1
Lambda_4_factor[8:14,3] <- 1
Lambda_4_factor[15:20,4] <- 1
```

Second, we define the alpha level and set the seed.

Third, we define the LNMs.
```{r LNM define models, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
# Define models
lnmMod_8_factor <-  lnm(df_factor, vars=vars_8,  lambda = Lambda_8_factor,  latents = Latents_8_factor, identification = "variance")
lnmMod_7_factor <-  lnm(df_factor, vars=vars_8,  lambda = Lambda_7_factor,  latents = Latents_7_factor, identification = "variance")
lnmMod_4_factor <-  lnm(df_factor, vars=vars_8,  lambda = Lambda_4_factor,  latents = Latents_4_factor, identification = "variance")
```

Fourth, we run the LNM.
```{r LNM run models, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
# Run & Remove non-sig latent edges:
lnmMod_8_prune <- lnmMod_8_factor %>% runmodel %>% psychonetrics::prune(alpha = alpha) 
lnmMod_7_prune <- lnmMod_7_factor %>% runmodel %>% psychonetrics::prune(alpha = alpha) 
lnmMod_4_prune <- lnmMod_4_factor %>% runmodel %>% psychonetrics::prune(alpha = alpha) 
```


Fit measures
```{r LNM fit models, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
lnmMod_8_prune@fitmeasures[c("chisq","pvalue","df", "rmsea","cfi", "tli", "aic.ll", "bic")]
lnmMod_7_prune@fitmeasures[c("chisq","pvalue","df", "rmsea","cfi", "tli", "aic.ll", "bic")]
lnmMod_4_prune@fitmeasures[c("chisq","pvalue","df", "rmsea","cfi", "tli", "aic.ll", "bic")]
```

Plot
```{r LNM Plots, eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
qgraph(lnmMod_8_prune@modelmatrices[[1]]$omega_zeta, labels = Latents_8_factor,
       theme = "colorblind", vsize = 10)
qgraph(lnmMod_7_prune@modelmatrices[[1]]$omega_zeta, labels = Latents_7_factor,
       theme = "colorblind", vsize = 10)
qgraph(lnmMod_4_prune@modelmatrices[[1]]$omega_zeta, labels = Latents_4_factor,
       theme = "colorblind", vsize = 10)
```




### Total score evaluation - using simple network analysis
- Not latent models
```{r eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
# build cluster scores
df3 <- df_factor %>% mutate(
  RE1 = PCL1 + PCL2 + PCL3,
  RE2 = PCL4 + PCL5,
  Av = PCL6 + PCL7,
  Na = PCL8 + PCL9 + PCL10 + PCL11,
  An = PCL12 + PCL13 + PCL14,
  Eb = PCL15 + PCL16,
  Aa = PCL17 + PCL18,
  Da = PCL19 + PCL20,

)

# bootnet

df4 <- df3 %>% select(RE1, RE2, Av, Na, An, Eb, Aa, Da, totalRTQ, totalGAD) 

net <- estimateNetwork(df4[,2:11], "ggmModSelect")
plot(net)
```

Now with ASI:

```{r eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
df5 <- df3 %>% select(RE1, RE2, Av, Na, An, Eb, Aa, Da, totalRTQ, totalASI)
netASI <-  estimateNetwork(df5[,2:11],default = "ggmModSelect") 
plot(netASI)
```

## Seperating those who chose emotional Pain and those who said fearful reaction
First, how many in each group
```{r eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
df %>% mutate(Q47 = factor(Q47, levels=c(1:2), labels = c("Emotional Pain","Fearful Reaction"))) %>% select(Q47)  %>%
  tbl_summary(label = list(Q47 ~ "Education"))
  

```
### Emotional Pain
```{r eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
# emotional pain first
df6 <- df3 %>% filter(Q47==1) %>% select(RE1, RE2, Av, Na, An, Eb, Aa, Da, totalRTQ, totalGAD)
net_Em_Pain <-  estimateNetwork(df6[,2:11],default = "ggmModSelect") 
plot(net_Em_Pain)
```
### Fearful Reaction
```{r eval=FALSE, message=TRUE, warning=FALSE, include=FALSE}
# emotional pain first
df7 <- df3 %>% filter(Q47==2) %>% select(RE1, RE2, Av, Na, An, Eb, Aa, Da, totalRTQ, totalGAD)
net_Fear <-  estimateNetwork(df7[,2:11],default = "ggmModSelect") 
plot(net_Fear)
```












# B & I
## 1. Stratification
Individual items
Q42 Having strong negative feelings such as fear or horror?	
Q43 Having strong negative feelings such as guilt, or shame?	
Q44 Having strong negative feelings of anger?

Select new items related to emotions and some relevant columns,convert all columns to numeric type and remove rows with NA values
```{r}
df4 <- df3 %>% 
  dplyr::select(Q42:Q44,
                Q2:Q21, #PCL
                Q37_1:Q37_9, #PHQ
                Q38_1:Q38_7, #GAD
                Q51_1:Q51_16, #ASI
                totalPCL, totalPHQ, totalGAD, totalASI, totalRTQ) %>% 
  mutate_all(as.numeric) %>% 
  na.omit()
```

Assess the correlation between the different emotions
```{r}
df5 <- df4 %>% 
  dplyr::select(Q42:Q44)

cor_matrix <- cor(df5)
round(cor_matrix, 2)

# Create a data frame from the matrix
df_plot <- as.data.frame(df5)

# Create a jitter plot matrix with correlation line
ggplot(df_plot, aes(x = Q42, y = Q43)) + 
  geom_jitter() +
  stat_smooth(method = "lm", se = FALSE, color = "red") + # Add a linear regression line
  labs(x = "Q42", y = "Q43") +
  theme_classic()

ggplot(df_plot, aes(x = Q42, y = Q44)) + 
  geom_jitter() +
  stat_smooth(method = "lm", se = FALSE, color = "red") + # Add a linear regression line
  labs(x = "Q42", y = "Q44") +
  theme_classic()

ggplot(df_plot, aes(x = Q43, y = Q44)) + 
  geom_jitter() +
  stat_smooth(method = "lm", se = FALSE, color = "red") + # Add a linear regression line
  labs(x = "Q43", y = "Q44") +
  theme_classic()
```

### 1.1 PTSD Internal structre
Analyze the relationship between the new emotion items and the other PTSD items using correlation and network analysis.

Correlation Analysis
```{r}
df6 <- df4 %>%   
  select(Q2:Q21, Q42, Q43, Q44) 

cor_matrix <- cor(df6, use = "pairwise.complete.obs")
print(cor_matrix, digits = 2)
```

Network Analysis
```{r echo=FALSE, message=FALSE, warning=FALSE}
net_PCL_new_items <- estimateNetwork(df6, "ggmModSelect")
plot(net_PCL_new_items, layout = "circle")
```

Exploratory Factor (EFA)
Assess whether the new emotion items align with the DSM-5 model by conducting EFA assuming a 4 factor structure.
```{r}
# Perform EFA using principal axis factoring (PAF) with oblimin rotation
efa_result <- fa(cor_matrix, nfactors = 4, rotate = "oblimin", fm = "pa", 
                 missing = "pairwise", scores = "regression")

# Print the factor loadings
print(efa_result$loadings, digits = 2, cutoff = 0.4)
```




Assess how the are related to the 8-factor model [not run because it exceeds running time]
```{r eval=FALSE, include=FALSE}
# Define
Latents_8_factor_new_items<-c("Re1","Re2", "Av", "Na", "An","Eb","Aa", "Da", "Fear", "Shame", "Anger")

# Variables
vars_all = c("Q2", "Q3", "Q4","Q5","Q6","Q7","Q8","Q9" ,"Q10",
        "Q11","Q12","Q13","Q14","Q15","Q16","Q17","Q18", "Q19","Q20","Q21", "Q42", "Q43", "Q44")

# Individual factors
Lambda_10_factors <- matrix(0,23,11)
Lambda_10_factors[1:3,1] <- 1
Lambda_10_factors[4:5,2] <- 1
Lambda_10_factors[6:7,3] <- 1
Lambda_10_factors[8:11,4] <- 1
Lambda_10_factors[12:14, 5] <- 1
Lambda_10_factors[15:16,6] <- 1
Lambda_10_factors[17:18,7] <- 1
Lambda_10_factors[19:20,8] <- 1
Lambda_10_factors[21,9] <- 1
Lambda_10_factors[22,10] <- 1
Lambda_10_factors[23,11] <- 1

# Run
lnmMod_10_factor <-  lnm(df6, vars=vars_all,  lambda = Lambda_10_factors,  latents = Latents_8_factor_new_items, identification = "variance")

# Prune
lnmMod_10_prune <- lnmMod_10_factor %>% runmodel %>% psychonetrics::prune(alpha = alpha) 

qgraph(lnmMod_10_prune@modelmatrices[[1]]$omega_zeta, labels = Latents_8_factor_new_items,
       theme = "colorblind", vsize = 10)

# Results
lnmMod_10_prune@fitmeasures[c("chisq","pvalue","df", "rmsea","cfi", "tli", "aic.ll", "bic")]
```

Using ordinary networks
- not using LNM (due to power)
- need to update the code
```{r}
# build factors scores
df4_NW <- df4 %>% 
  mutate(
    RE1 = Q2 + Q3 + Q4,
    RE2 = Q5 + Q6,
    Av = Q7 + Q8,
    Na = Q9 + Q10 + Q11 + Q12,
    An = Q13 + Q14 + Q15,
    Eb = Q16 + Q17,
    Aa = Q18 + Q19,
    Da = Q20 + Q21) %>% 
  select(RE1, RE2, Av, Na, An, Eb, Aa, Da, Q42, Q43, Q44)

# Estimate network
net <- estimateNetwork(df4_NW, "ggmModSelect")
plot(net, layout = "circle")
```

## 1.2. EXTERNAL
Full data matrix for the emotions and the scale total scores
```{r}
df7 <- df4 %>% 
  dplyr::select(Q42:Q44, totalPCL, totalPHQ, totalGAD, totalASI, totalRTQ)

cor_matrix <- cor(df7)
round(cor_matrix, 2)
```


Adjusting for PTSD overall
Linear regression models are created to assess the association between each emotion item and the four scale total scores while adjusting for PTSD overall (measured by totalPCL). The lm function is used to fit the models, and the summary function is used to print the model summary.

1. Rumination - RTQ
```{r}
model_rtq <- lm(totalRTQ ~ totalPCL + Q42 + Q43 + Q44, data = df7)

# Print the model summary
summary(model_rtq)
```

2. Anxiety - ASI
```{r}
model_asi <- lm(totalASI ~ totalPCL + Q42 + Q43 + Q44, data = df7)

# Print the model summary
summary(model_asi)
```

3. Anxiety - GAD
```{r}
model_gad <- lm(totalGAD ~ totalPCL + Q42 + Q43 + Q44, data = df7)

# Print the model summary
summary(model_gad)
```

4. Depression - PHQ
```{r}
model_phq <- lm(totalPHQ ~ totalPCL + Q42 + Q43 + Q44, data = df7)

# Print the model summary
summary(model_phq)
```


## Network annalysis
To estimate the network, the provided code selects the relevant variables using the select function, estimates the network using the estimateNetwork function, and plots the network using the plot function with a circular layout.
```{r}
df8 <- df7 %>%   
  select(totalPCL, totalPHQ, totalGAD, totalASI, totalRTQ, Q42, Q43, Q44) 

net <- estimateNetwork(df8, "ggmModSelect")
plot(net, layout = "circle")

```



### 2. INDIVIDUALS
### 2.1 Internal
#### Categorical
Calculate how many individuals report high (>=2 on a 0-4 scale) fear, shame, and anger
```{r}
df6 <- df4 %>% 
  select(Q42:Q44) %>% 
  mutate(across(.cols = everything(), as.numeric)) %>% 
  mutate(Q42_low_high = ifelse(Q42 < 3, "low", "high"),
         Q43_low_high = ifelse(Q43 < 3, "low", "high"),
         Q44_low_high = ifelse(Q44 < 3, "low", "high"))

df_summary <- df6 %>%
  select(Q42_low_high, Q43_low_high, Q44_low_high) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  group_by(variable, value) %>%
  summarise(n = n(), percentage = n()/nrow(df4)*100)
```
The distribution is not even. More individuals report high shame than fear or anger. 

Next, we test how many individuals with high fear report high/low shame or high/low anger. We do this for all three emotions.
```{r}
# Calculate the number and percentage of individuals who are high or low on other variables within the Q42 high and low subsets
df6 %>%
  group_by(Q42_low_high) %>%
  summarise(Q43_high = sum(Q43 >= 3),
            Q43_low = sum(Q43 < 3),
            Q44_high = sum(Q44 >= 3),
            Q44_low = sum(Q44 < 3),
            percentage_Q43_high = Q43_high/n()*100,
            percentage_Q43_low = Q43_low/n()*100,
            percentage_Q44_high = Q44_high/n()*100,
            percentage_Q44_low = Q44_low/n()*100) %>%
  print()

df6 %>%
  group_by(Q43_low_high) %>%
  summarise(Q42_high = sum(Q42 >= 3),
            Q42_low = sum(Q42 < 3),
            Q44_high = sum(Q44 >= 3),
            Q44_low = sum(Q44 < 3),
            percentage_Q42_high = Q42_high/n()*100,
            percentage_Q42_low = Q42_low/n()*100,
            percentage_Q44_high = Q44_high/n()*100,
            percentage_Q44_low = Q44_low/n()*100) %>%
  print()


df6 %>%
  group_by(Q44_low_high) %>%
  summarise(Q42_high = sum(Q42 >= 3),
            Q42_low = sum(Q42 < 3),
            Q43_high = sum(Q43 >= 3),
            Q43_low = sum(Q43 < 3),
            percentage_Q42_high = Q42_high/n()*100,
            percentage_Q42_low = Q42_low/n()*100,
            percentage_Q43_high = Q43_high/n()*100,
            percentage_Q43_low = Q43_low/n()*100) %>%
  print()
```
There are relevant differences between individuals with high fear or high shame. 

Next, we want to assess, whether individuals who have predomninately fear vs shame do report differences in their levels of psychopathology (i.e., can we separate two subgroups based on fear vs shame).
```{r}
# Add a variable indicating two groups based on Q42 and Q43
df9 <- df4 %>%
  mutate(Q42_low_high = ifelse(Q42 < 3, "low", "high"),
         Q43_low_high = ifelse(Q43 < 3, "low", "high"),
         Q44_low_high = ifelse(Q44 < 3, "low", "high")) %>% 
  mutate(Q42_Q43_group = ifelse(Q42_low_high == "high" & Q43_low_high == "low", "Q42_high_Q43_low",
                                ifelse(Q42_low_high == "low" & Q43_low_high == "high", "Q42_low_Q43_high", NA)))

# Summarize the mean and standard deviation (or median) of totalPCL, totalPHQ, totalGAD, totalASI, and totalRTQ by Q42_Q43_group
df9 %>%
  group_by(Q42_Q43_group) %>%
  summarise(median_totalPCL = median(totalPCL),
            median_totalPHQ = median(totalPHQ),
            median_totalGAD = median(totalGAD),
            median_totalASI = median(totalASI),
            median_totalRTQ = median(totalRTQ))
```
There are no relevant differences.

### Cluster Analysis
We use cluster analysis to check for subgroups. We define three clusters (due to three emotions). First, we use the three emotions and all relevant constructs (PCL, RTW, GAD, ASI, PHQ)
```{r}
# Select the relevant columns from the dataframe
df_subset <- df7

# Use hierarchical clustering to determine the number of clusters
fviz_nbclust(df_subset, hcut, method = "wss", k.max = 10)

# Perform K-means clustering with the recommended number of clusters
cluster_result <- kmeans(df_subset, centers = 3)

# Print the cluster centers
print(cluster_result$centers)

# Add cluster assignments to the original dataframe
df_clusters <- df7 %>%
  mutate(cluster = cluster_result$cluster)

# Summarize the cluster assignments by variable
df_summary <- df_clusters %>%
  group_by(cluster) %>%
  summarise_all(funs(mean))
print(df_summary)
```
The separation is along severity and not different symptoms/psychopathologies.

Focusing on PTSD only.
```{r}
# Select the relevant columns from the dataframe
df_subset_2 <- df7 %>% 
  dplyr::select(totalPCL, Q42,Q43, Q44) %>% 
  filter(totalPCL >= 33)

# Check for missing values
sum(is.na(df_subset_2))

# Standardize the data
df_subset_std <- scale(df_subset_2)

# Perform clustering using CLARA algorithm
clara_result <- clara(df_subset_std, k = 3, metric = "euclidean", samples = 10)

# Add cluster assignments to the original dataframe
df_clusters <- df_subset_2 %>%
  mutate(cluster = clara_result$clustering)

# Summarize the cluster assignments by variable
df_summary <- df_clusters %>%
  group_by(cluster) %>%
  summarise_all(funs(mean))
print(df_summary)
```
Still based on severity.

Could use PCA....results are still based on severity.
